---
title: Introduction to NAU
author: ''
date: ''
slug: []
categories: []
tags: []
---
<style>
table {
    margin: 25px 0;
    font-family: sans-serif;
    text-align: center;
    table-layout: fixed;
    width: 500px;
}
th {
    font-weight: bold;
    padding: 10px;
    text-shadow: 1px 1px 1px #fff;
    text-align: center;
}
td {
    background-color: #7FFFD4;
    border: solid 1px #DDEEEE;
    padding: 10px;
    text-shadow: 1px 1px 1px #fff;
}

</style>
<body>

## Utility is Noisy

```{r include=F}
library(ggplot2)
library(tidyr)
library(dplyr)
library(gridExtra)
library(ggrepel)
library(cowplot)
library(ellipse)
cbp1 = c("#999999", "#E69F00", "#56B4E9", "#009E73",
         "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

signal.i = 10
a=1  # sigma i
b=1  # sigma j
c=1  # sigma c
sigma.c = sqrt(c^2 + b^2)
    
#min.val and max.val for which values of x to calculate
min.val = qnorm(0.00001, mean = signal.i, sd=(sqrt(a^2 + sigma.c^2)))
max.val = qnorm(0.99999, mean = signal.i, sd=(sqrt(a^2 + sigma.c^2)))
distr.table = data.frame(x.d = seq(from = min.val, to = max.val, length.out = 500))

#compute the prior distribution
distr.table$prior.value = sapply(distr.table$x.d, FUN = function(x){dnorm(x, mean = signal.i, sd=a)})

#data for ggplot
plot.what = c("Prior Value" = "prior.value")
col.select = which(colnames(distr.table) %in% plot.what)
LD = distr.table %>% gather(key = distributions, value = density, col.select)
    
#plot graph
plot1 = ggplot(data=LD, aes(x=x.d, y= density)) + 
        geom_area(alpha=0.2, position = "identity", aes(fill=distributions)) +
        geom_line(aes(colour=distributions)) +
        lims(x=c(5, 15))+
        scale_colour_manual(name = NULL,
                            breaks=c("prior.value"),
                            values = c("prior.value" = cbp1[4]),
                            labels = c("Prior Value")) +
        scale_fill_manual(name = NULL,
                          breaks=c("prior.value"),
                          values = c("prior.value" = cbp1[4]),
                          labels = c("Prior Value")) +
        theme_classic() + 
        labs(title=NULL, x="Utility", y="Probability Density") +
        theme(legend.justification=c(1,1), legend.position="none")
   #   if(do.posterior){plot1 = plot1 + geom_vline(aes(xintercept = E.posterior), colour = cbp1[2], show.legend = T, linetype = "dashed")} 
  #    if(do.threshold){plot1 = plot1 + geom_vline(aes(xintercept = input$t), show.legend = F)}
```
```{r echo=F}
suppressWarnings(print(plot1))
```

This means that utility of an action is not known exactly. Instead, there is a probability distribution. This allows us to model the uncertainty over how much a person likes a particular good.

## Affiliated Utility

```{r include=F}
do.uncertainty = F
do.perspective.1 = F
numgoods = 60
    
## function for initial plots
create.plot = function(title, plot.table){#plot.table = plot.table.aff; title = "Affiliated Value"
  #create plot
  if(do.perspective.1){plot.table$u2 = plot.table$u1}
  ggplot(data=plot.table, aes(x=u1, y=u2, label=label)) +
    geom_point(data = plot.table[plot.table$label == "",], color = "grey50") +
    geom_point(data = plot.table[plot.table$label != "",], color = "red") + 
    geom_text_repel(box.padding = 0.5, nudge_x = .75, nudge_y = 0) +
    labs(title=title, x="Utility Person 1", y="Utility Person 2") +
    lims(x=c(0, 10),
       y= c(0, 10)) +
    theme_classic()
}

#calculating ellipses function
    ellipse.df = function(df, dis){
      df_ell <- data.frame()
      cor = 2 / (dis + 2)
      for(n in 1:nrow(df)){ #n = 1
        if(do.perspective.1){
          sigma.u2.unc = sqrt(2*dis^2 + b^2)
          mean.u2 = df$u1[n] 
        } else {
          sigma.u2.unc = b
          mean.u2 = df$u2[n]
        }
        df_ell <- rbind(df_ell, cbind(as.data.frame(ellipse(cor, 
                                                            scale=c(a, sigma.u2.unc), 
                                                            centre=c(df$u1[n], mean.u2))), datpoint=n))
      }
    df_ell$datpoint = as.factor(df_ell$datpoint)
    return(df_ell)
    }
    
    plot.table = data.frame(x.d = seq(from = 0, to = 10, length.out = numgoods))
    plot.table$base = runif(n = numgoods, min = plot.table$x.d[1], max = tail(plot.table$x.d, 1))
    # include choc and bloc points
    choc = which.min(abs(plot.table$base - quantile(plot.table$base, 0.75)))
    broc = which.min(abs(plot.table$base - quantile(plot.table$base, 0.25)))
    ix_label <- c(broc, choc)
    plot.table$label = ""
    plot.table$label[ix_label] <- c("Broccoli", "Chocolate Cake")
    
    
    plot.table.com = plot.table
    plot.table.com$u1 = plot.table.com$base
    plot.table.com$u2 = plot.table.com$base
    ComVal = create.plot("Common Value", plot.table.com)
    
    plot.table.aff = plot.table
    plot.table.aff$u1 = sapply(plot.table.aff$base, FUN = function(x){rnorm(n=1, mean = x, sd= c)})
    plot.table.aff$u2 = sapply(plot.table.aff$base, FUN = function(x){rnorm(n=1, mean = x, sd= c)})
    AffVal = create.plot("Affiliated Value", plot.table.aff)
    AffVal2 = create.plot("", plot.table.aff)
    
    plot.table.ind = plot.table
    plot.table.ind$u1 = plot.table.ind$base
    plot.table.ind$u2 = runif(n = numgoods, min = plot.table.ind$x.d[1], max = tail(plot.table.ind$x.d, 1))
    IndVal = create.plot("Independent Value", plot.table.ind)
    
    if(do.uncertainty){
      #drawing
      df_ell = ellipse.df(plot.table.com, dis = 0.001) 
      ComVal =  ComVal + geom_polygon(data=df_ell, aes(x=x, y=y, group=datpoint, label = NA), size=1, alpha = 0.2, fill = "grey50")
      df_ell = ellipse.df(plot.table.aff, dis = c) 
      AffVal = AffVal + geom_polygon(data=df_ell, aes(x=x, y=y, group=datpoint,  label = NA), size=1, alpha = 0.2, fill = "grey50")
      df_ell = ellipse.df(plot.table.ind, dis = Inf) 
      IndVal = IndVal + geom_polygon(data=df_ell, aes(x=x, y=y, group=datpoint,  label = NA), size=1, alpha = 0.2, fill = "grey50")
    }
    
    plots = cowplot::plot_grid(
      ComVal, AffVal, IndVal, nrow = 1)
  
```
```{r echo=F}
 suppressWarnings(print(AffVal2))
```

This means that two people's utilities are interdependent. If person 1 likes a particular good, then probably person 2 likes it too. The extreme version of affiliation is common value where two people's preferences can be represented by the same utility function.

## Things Only Get Interesting If You Put the Two Together

<table>
  <tr>
    <th style = "border: solid 0px #DDEEEE;"></th>
    <th>Simple</th>
    <th>Noisy</th>
  </tr>
  <tr>
    <th>Affiliated</th>
    <td>Not Interesting</td>
    <td style = "background-color: #FFE4C4;" >Interesting</td>
  </tr>
  <tr>
    <th>Independent</th>
    <td>Not Interesting</td>
    <td>Not Interesting</td>
  </tr>
</table>

If Utility is Noisy *and* Affiliated,  
$\rightarrow$ then there is a gap between Willingness to Pay and Willingness to Accept  
$\rightarrow$ and fundamental theorems of welfare do not apply  
$\rightarrow$ we enter a world of social learning and Bayesian updating 

# Focus on WTP-WTA Gap

Intuition: The fact that someone wants to sell a good sends a signal about its worth.


```{tikz, echo=F, fig.cap = 'A WTP-WTA gap exists in the red zone', fig.ext = 'png', out.width = "75%"}
\definecolor{badcol}{HTML}{FFE4C4}
\definecolor{goodcol}{HTML}{7FFFD4}
\begin{tikzpicture}
\fill [badcol] (0,0) rectangle (4.9,5);
\draw[help lines, color=gray!30, dashed] (0,0) grid (4.9,4.9);
\draw[->,ultra thick, color=goodcol] node[below, color=black]{\tiny Certainty} (-0.01,0)--node[below, color=black] {Noise in Utility} (5,0) node[right, color=black]{\tiny No Info.};
\draw[-,ultra thick, color=goodcol] node[left, color=black]{\tiny Independent} (0,0)--node[left, color=black] {Affiliation} (0,5) node[left, color=black]{\tiny Common Value};
\end{tikzpicture}
```

On the y-axis (utility has zero noise) there is no gap and the fundamental theorems of welfare apply. On the x-axis (utility has zero affiliation) there is also no gap and the fundamental theorems of welfare apply ex-ante^[Ex-ante pareto efficiency in the sense that no person can increase expected utility without someone else decreasing expected utility. However, if the underlying utility function were known, then a pareto superior outcome could be found.]. Moving up the y-axis means that there are fewer potential gains from trading, and moving along the x-axis implies those gains are harder to find. 

However, the interaction of Noise and Affiliation opens up a WTP-WTA gap and the further you go from the axes, the bigger the gap becomes. 


</body>